---
title: "Models"
author: "Patrick Breheny"
---

```{r setup, include=FALSE}
library(ncvreg)
knitr::opts_knit$set(aliases=c(h = 'fig.height', w = 'fig.width'))
knitr::opts_chunk$set(comment="#", collapse=TRUE, cache=FALSE, tidy=FALSE)
knitr::knit_hooks$set(small.mar = function(before, options, envir) {
  if (before) par(mar = c(4, 4, .1, .1))
})
```

`ncvreg` fits models that fall into the penalized likelihood framework.  Rather than estimating $\bb$ by maximizing the likelihood, in this framework we estimate $\bb$ by minimizing the objective function
$$
Q(\bb|\X, \y) = \frac{1}{n}L(\bb|\X,\y) + P_\lam(\bb),
$$
where $L(\bb|\X,\y)$ is the loss (typically, $-2$ times the log-likelihood), $P_\lam(\bb)$ is the penalty, and $\lam$ is a regularization parameter that controls the tradeoff between the two components.  This article describes the different loss models available in `ncvreg`; see [penalties](penalties.html) for more information on the different penalties available.

# Gaussian (linear regression)

In linear regression, the loss function is simply the squared error loss:
$$
L(\bb|\X,\y) = \norm{\y-\X\bb}_2^2
$$
this loss is proportional to $-2 \times$ log-likelihood if the outcome $\y$ follows a normal distribution with constant variance and mean given by $\X\bb$.

In the `Prostate` data packaged with `ncvreg`, the response is the prostate specific antigen (PSA), measured on the log scale, and follows an approximate normal distribution; see `?Prostate` for more information on the data set.  Loading this data set into R,

```{r}
data(Prostate)
X <- Prostate$X
y <- Prostate$y
```

By default, `ncvreg` fits a linear regression model with a minimax concave penalty (MCP):

```{r}
fit <- ncvreg(X, y)
```

This produces a path of coefficient estimates, which we can plot with

```{r fit-linear, h=4, w=6, small.mar=TRUE}
plot(fit)
```

Although the least squares loss function is convex, the MCP penalty is not.  The resulting objective function, therefore, may or may not be convex.  `ncvreg`uses a local convexity diagnostic, as described in [Breheny and Huang (2011)](http://myweb.uiowa.edu/pbreheny/pdf/Breheny2011.pdf), to identify the regions of the coefficient path where the objective function is not convex; this is the gray shaded region in the plot.  Users should be aware that solutions in this region may only be local optima of the objective function, not global ones.

Post-selection inference is available using the `summary` method:

```{r}
summary(fit, lambda=0.05)
```

The local marginal false discovery rate (mfdr) is given for each of the selected features.  Roughly, this corresponds to the probability that the given feature is marginally independent of the residuals at that value of $\lam$.  In this case, it would appear that `lcavol`, `svi`, and `lweight` are clearly associated with the response, even after adjusting for the other variables in the model, while `lbph`, `age`, and `pgg45` may be false positives selected simply by chance.  For more information on `summary()` and its various options, see [here](inference-local.html).

Typically, one would carry out cross-validation for the purposes of assessing the predictive accuracy of the model at various values of $\lambda$:

```{r, h=5, w=6}
cvfit <- cv.ncvreg(X, y)
plot(cvfit)
```

By default, the cross-validation error (CVE) is plotted; for all models in `ncvreg`, the cross-validation error is defined as
$$
\begin{align*}
\CVE(\lam) &= \frac{1}{n} \sum_i L\{y_i, \eta_{-i}(\lam)\}\\
\eta_{-i}(\lam) &= \sum_j x_{ij}\bh_j(-i,\lam),
\end{align*}
$$
where $\bbh(-i,\lam)$ denotes the estimated regression coefficients at $\lam$ for the fold in which observation $i$ has been left out.  The loss function is determined by the type of model; for least squares loss, therefore,
$$
\CVE(\lam) = \frac{1}{n} \sum_i\{y_i-\eta_{-i}(\lam)\}^2\\
$$

Alternatively, one can plot $\hat{\sigma}(\lam)$, the signal-to-noise ration (SNR), or $R^2$:
```{r, h=6, w=6}
par(mfrow=c(2,2))
plot(cvfit, type='cve')
plot(cvfit, type='scale')  # sigma hat
plot(cvfit, type='snr')
plot(cvfit, type='rsq')
```

Calling `summary` on a `cv.ncvreg` object will provide a summary of these quantities at the value which minimizes $\CVE$:

```{r}
summary(cvfit)
```

To access the elements of the fit, `coef` and `predict` methods are provided.  For example, `coef(fit, lambda=0.02)` returns the estimated coefficients at $\lambda$=0.02, while `coef(cvfit)` returns the estimated coefficients at the value of $\lam$ minimizing CVE.

# Binomial (logistic regression)

In logistic regression, the loss function is:
$$
L(\bb|\X,\y) = -2\sum_{i:y_i=1}\log\ph_i - 2\sum_{i:y_i=0}\log(1-\ph_i);
$$
this loss is $-2 \times$ log-likelihood if the outcome $\y$ follows a binomial distribution with probabilities given by:
$$
\ph_i = \frac{\exp(\eta_i)}{1+\eta_i},
$$
where $\be = \X\bb$ denotes the linear predictors.  The `Heart` data provides an example of data that can be used with logistic regression.  Loading this data set into R,

```{r}
data(Heart)
X <- Heart$X
y <- Heart$y
```

One can change the loss function by specifying `family`; to fit a penalized logistic regression model,

```{r}
fit <- ncvreg(X, y, family='binomial')
```

As before, you can call `plot`, `coef`, `predict`, `summary`, etc. on `fit`:

```{r}
summary(fit, lambda=0.02)
```

Cross-validation is similar, although (a) there is a new option, `type='pred'` for cross-validated prediction error (misclassification error) and (b) `type='scale'` is no longer an option:

```{r, h=6, w=6}
cvfit <- cv.ncvreg(X, y, family='binomial')
par(mfrow=c(2,2))
plot(cvfit, type='all')
```

Note that, as defined above, cross-validation error is the cross-validated deviance.  At its optmium, the penalized logistic regression model can predict about 73\% of coronary heart disease cases correctly (27\% misclassification).

# Poisson

In Poisson regression, the loss function is:
$$
L(\bb|\X,\y) = 2\sum_i \left\{\mh_i - y_i\log\mh_i\right\};
$$
this loss is $-2 \times$ log-likelihood if the outcome $\y$ follows a Poisson distribution with rate parameter given by:
$$
\mh_i = \exp(\eta_i).
$$
To fit a penalized Poisson regression model with `ncvreg`:
```{r, eval=FALSE}
fit <- ncvreg(X, y, family='poisson')
```

# Cox proportional hazards

The above models all fall into the category of distributions known as exponential families (hence the `family`) argument.  `ncvreg` also allows users to fit Cox proportional hazards models, although these models fall outside this framework and are therefore fit using a different function, `ncvsurv`.  In Cox regression, the loss function ($-2 \times$ the partial log-likelihood) is
$$
L(\bb|\X,\y) = -2\sum_{j=1}^{m} d_j \eta_j + 2\sum_{j=1}^{m} d_j \log\left\{\sum_{i \in R_j} \exp(\eta_i)\right\},
$$
where $t_1 < t_2 < \ldots < t_m$ denotes an increasing list of unique failure times indexed by $j$ and $R_j$ denotes the set of observations still at risk at time $t_j$, known as the risk set.

The `Lung` data (see `?Lung` for more details) provides an example of time-to-event data that can be used with Cox regression.  Loading this data set into R,

```{r}
data(Lung)
X <- Lung$X
y <- Lung$y
```

To fit a penalized Cox regression model,

```{r}
fit <- ncvsurv(X, y)
```

As before, you can call `plot`, `coef`, `predict`, `summary`, etc. on `fit`:

```{r}
summary(fit, lambda=0.02)
```

Cross-validation is similar:

```{r models-cox-cv, h=3.5, w=6}
cvfit <- cv.ncvsurv(X, y)
par(mfrow=c(1,2))
plot(cvfit, type='cve')
plot(cvfit, type='rsq')
```

In addition to the quantities like coefficients and number of nonzero coefficients that `predict` returns for other types of models, `predict()` for an `ncvsurv` object can also estimate the baseline hazard (using the Kalbfleish-Prentice method) and therefore, the survival function.  A method to plot the resulting function is also available:

```{r models-cox-surv-single, h=4, w=4}
S <- predict(fit, X[1,], type='survival', lambda=0.02)
S(365)   # Estiamted survival at 1 year
plot(S, xlim=c(0,200))
```

When multiple subjects are involved in the prediction:

```{r models-cox-surv-multiple, h=4, w=4}
S <- predict(fit, X, type='survival', lambda=0.02)
S[[1]](365)  # Estimated survival at 1 year for subject 1
S[[2]](365)  # Estimated survival at 1 year for subject 2
plot(S, xlim=c(0,200))
```
