---
title: "Penalties"
author: "Patrick Breheny"
---

```{r setup, include=FALSE}
library(ncvreg)
knitr::opts_knit$set(aliases=c(h = 'fig.height', w = 'fig.width'))
knitr::opts_chunk$set(comment="#", collapse=TRUE, cache=FALSE, tidy=FALSE)
knitr::knit_hooks$set(small.mar = function(before, options, envir) {
  if (before) par(mar = c(4, 4, .1, .1))
})
```

<div style='text-align: center; font-size: 200%; font-weight: 700; margin: 20pt;'><i class="fa fa-wrench"></i> Under construction <i class="fa fa-wrench"></div>

`ncvreg` fits models that fall into the penalized likelihood framework.  Rather than estimating $\bb$ by maximizing the likelihood, in this framework we estimate $\bb$ by minimizing the objective function

$$
Q(\bb|\X, \y) = L(\bb|\X,\y) + P_\lam(\bb),
$$

where $L(\bb|\X,\y)$ is the loss (deviance) and $P_\lam(\bb)$ is the penalty.  This article describes the different penalties available in `ncvreg`; see [models](models.html) for more information on the different loss functions available.  Throughout, linear regression and the `Prostate` data set is used

```{r}
data(Prostate)
X <- Prostate$X
y <- Prostate$y
```

# MCP

This is the default penalty in `ncvreg`.

EQUATION HERE.

PICTURE OF PENALTY.

The following figure illustrates the effect of changing $\gamma$:

```{r penalty-mcp-gamma, h=3.5, w=8}
par(mfrow=c(1,3))
fit <- ncvreg(X, y, gamma=1.5)
plot(fit, main=expression(paste(gamma,"=",1.5)))
fit <- ncvreg(X, y)
plot(fit, main=expression(paste(gamma,"=",3)))
fit <- ncvreg(X, y, gamma=10)
plot(fit, main=expression(paste(gamma,"=",10)))
```

At smaller $\gamma$ values, the estimates transition rapidly from 0 to their unpenalized solutions; this transition happens more slowly and gradually at larger $\gamma$ values.  Note that one consquence of these rapid transitions at low $\gamma$ values is that the solutions are less stable (the gray region depicting the region of the solution path that is not locally convex is larger).

# SCAD

EQUATION HERE.

PICTURE OF PENALTY.

# Lasso

EQUATION HERE.

PICTURE OF PENALTY.

```{r penalty-comp, h=3.5, w=8}
par(mfrow=c(1,3))
fit <- ncvreg(X, y)
plot(fit, main="MCP")
fit <- ncvreg(X, y, penalty="SCAD")
plot(fit, main="SCAD")
fit <- ncvreg(X, y, penalty="lasso")
plot(fit, main="Lasso")
```

# Elastic Net and MNet

EQUATION HERE.

```{r penalty-mnet, h=3.5, w=8}
par(mfrow=c(1,3))
fit <- ncvreg(X, y)
plot(fit, main=expression(paste(alpha,"=",1)))
fit <- ncvreg(X, y, alpha=0.5)
plot(fit, main=expression(paste(alpha,"=",0.5)))
fit <- ncvreg(X, y, alpha=0.1)
plot(fit, main=expression(paste(alpha,"=",0.1)))
```

MENTION ELASTIC NET
